---
title: "Homework 3"
author: "Autrin Peralta"
date: 03/08/2026
format: html
editor: visual
---

# Set Up

```{r}
#| label: reading in packages
#| message: false
library(tidyverse)#general use for data wrangling, cleaning, etc.
library(janitor)
library(calecopal)
library(readr)
salinity <- read_csv("~/ENVS-193DS/HW3/data/salinity-pickleweed (1).csv")


```

# Problem 1

a\. The two appropriate tests are the Pearson's correlation and the Spearman rank correlation. Pearson's is a parametric test that would assume that the salinity and California pickleweed biomass are both continuous variables with a normally distributed relationship measuring the strength of linearity. The Spearman is a non-parametric test that would measure the monotonic relationship between these two variables using ranks.

b\.

```{r}
#| label: Salinity vs biomass visualization
#| message: false
ggplot(data = salinity, #data set to plot
       aes(x = salinity_mS_cm, #set x-axis
           y = pickleweed, #set y-axis
           color = salinity_mS_cm)) +
  geom_point(color = "tan3") + #create scatterplot and channeg color of points
  theme_minimal() + #Put background as white
  labs(x = "Soil Salinity(mS/cm)", #name x-axis
       y = "Pickleweed Biomass(g)", #name y-axis
       title = "As Salinity Increases Biomass Increases") + #Make title
  theme(legend.position = "none") #remove legend
```

c\.

```{r}
#| label: Linear model function
#| message: false
pickleweed_model <- lm( #creating model object and creating linear model function
  pickleweed ~ salinity_mS_cm, #formula
  data = salinity #pulling data from "salinity"
)
```

```{r}
#| label: base R residuals
#| message: false
par(mfrow = c(1,2)) #display plots ina a 1x2 grid
plot(pickleweed_model) #plotting model data
```

The residual graphs show that the data has homoscedasicity and a clear normal distribution between the two variables . This means the Pearson's correlation test will be the most effective.

```{r}
cor.test(salinity$salinity_mS_cm,salinity$pickleweed, # doing a coorelation test pulling from salinity data set
         method = "pearson") # using Pearson's coorelation test
```

I checked fro normality and homoscedasticity in the qq plot and residual vs fitted plots. This would prove a normal distribution of the variables to justify using the Pearson's correlation test

d\.

For these variables we used the Pearson's coorelation test because both variables were continious and showed normal distirbution. We found a moderate relationship between soil salinity(mS/cm) and pickleweed biomass(g)(Pearson's r = 0.53, t(21) = 2.9, p = 0.01, ⍺ = 0.5). This shows that the correlation between salinity and biomass is significant which indicated that higher salinity soils relate to pickleweeds with larger biomass.

e\.

Larger, heavier pickleweeds have a clear positive relationship with soil salinity. This means to have the greatest success in repopulating the pickleweeds we should plant them in places with higher salinity.
