---
title: "Homework 3"
author: "Autrin Peralta"
date: 03/08/2026
format: html
editor: visual
---

# Set Up

```{r}
#| label: reading iun packages
#| message: false
library(tidyverse)#general use for data wrangling, cleaning, etc.
library(janitor)

library(readr)
salinity <- read_csv("~/ENVS-193DS/HW3/data/salinity-pickleweed (1).csv")


```

# Problem 1

a\. The two appropriate tests are the Pearson's correlation and the Spearman rank correlation. Pearson's is a parametric test that would assume that the salinity and California pickleweed biomass are both continuous variables with a normally distributed relationship measuring the strength of linearity. The Spearman is a non-parametric test that would measure the monotonic relationship between these two variables using ranks.

b\.

```{r}
#| label: Salinity vs biomass visualization
#| message: false
ggplot(data = salinity,
       aes(x = salinity_mS_cm,
           y = pickleweed,
           color = pickleweed)) + 
  geom_point(color = "forestgreen") + 
  theme_minimal() +
  labs(x = "Soil Salinity(mS/cm)",
       y = "Pickleweed Biomass(g)",
       title = "As Salinity Increases Biomass Increases") +
  theme(legend.position = "none")
```
