---
title: "Homework 3"
author: "Autrin Peralta"
date: 03/08/2026
format: html
editor: visual
---

# Set Up

```{r}
#| label: reading in packages
#| message: false
library(tidyverse)#general use for data wrangling, cleaning, etc.
library(janitor)
library(calecopal)
library(readr)
salinity <- read_csv("~/ENVS-193DS/HW3/data/salinity-pickleweed (1).csv")


```

# Problem 1

a\. The two appropriate tests are the Pearson's correlation and the Spearman rank correlation. Pearson's is a parametric test that would assume that the salinity and California pickleweed biomass are both continuous variables with a normally distributed relationship measuring the strength of linearity. The Spearman is a non-parametric test that would measure the monotonic relationship between these two variables using ranks.

b\.

```{r}
#| label: Salinity vs biomass visualization
#| message: false
ggplot(data = salinity, #data set to plot
       aes(x = salinity_mS_cm, #set x-axis
           y = pickleweed, #set y-axis
           color = salinity_mS_cm)) +
  geom_point(color = "tan3") + #create scatterplot and channeg color of points
  theme_minimal() + #Put background as white
  labs(x = "Soil Salinity(mS/cm)", #name x-axis
       y = "Pickleweed Biomass(g)", #name y-axis
       title = "As Salinity Increases Biomass Increases") + #Make title
  theme(legend.position = "none") #remove legend
```

c\.

```{r}
par(mfrow = c(1, 2))
qqnorm(salinity$salinity_mS_cm, main = "Salinity")

qqline(salinity$salinity_mS_cm, col = "red")

qqnorm(salinity$pickleweed, main = " Pickleweed Biomass")

qqline(salinity$pickleweed, col = "red")
```

Both QQ plots show a normal minimal skew from the line and only a few outliers showing that both variables are normally distributed. This means the Pearson test will be the most effective

d\.

```{r}
cor.test(salinity$salinity_mS_cm,salinity$pickleweed,
         method = "pearson")
```

We found a moderate relationship between soil salinity(mS/cm) and pickleweed biomass(g)(Pearson's r = 0.53, t(21) = 2.9, p = 0.01, ⍺ = 0.5). This shows that the coorelation between salinity and biomass is significant wihich indicated that higher salinity soils relate to pickleweeds with larger biomasses.
